### 2016-08-29 ReviewComments/review2.tex
Revised response
some more responses

\response{
We agree the fact that the gap parameter is an old concept. Nevertheless, our major contribution in
modeling lies in unifying existing movement patterns in the trajectory domain. 
The gap parameter is naturally introduced to 
solve the loose-connection anomaly (as in Figure 2).
%which is suffered by the state-of-the-art generalization (i.e., Platoon pattern). 
In the revision, we add Section 2.5 to accredit the origin of \emph{gap} concept
and distinguish the sequential mining problem with our problem. 
}

@@

\response{
We agree that if no other parameters are given, the maximum number
of patterns discovered could be as large as ${ \mathbb{O} \choose M}$, which may
be less interesting to users.
However, such a case rarely occurs due to two reasons. First,
in real trajectory mining, parameter $M$
is associated with temporal parameters $K,L,G$ and they jointly prune many false
patterns. Second, the nature of our solutions prefer 
a larger pattern when possible~\footnote{TRPM shrinks an invalid larger pattern until it is valid. SPARE grows a valid smaller pattern until it is about to be invalid. Therefore both algorithms tend to find larger patterns.}.
For example, if $(o_1,o_2,o_3)$ is 
a pattern and $M = 2$, 
then both TRPM and SPARE only output $(o_1,o_2,o_3)$ but not
any of its subsets. This effectively compresses smaller patterns.

In addition, when the temporal parameters $K,L,G$ are given,
a larger $M$ actually decreases the number of resulted patterns. 
This is because there are less patterns having more than $M$ objects. 
As shown in Figures 7 (a)(g)(m), a larger $M$ actually reduces the running time of both TRPM and SPARE.
}

@@

\response{We agree that domain-specific constraints
on spatiotemporal mining could be similar to GCMP.
However, these domain-specific constraints 
are nontrivial to be generalized. Since our GCMP model as well as solutions
do not leverage domain knowledge, we can support
pattern discovery in a broader context. We add the discussion in Section 2.3.
}

>>>>

\response{
We admit that the gap parameter is also used in
mining Gap-restricted Sequential Patterns (GSP). However, the novelty
of this paper is not merely on introducing the gap-constraint to the
co-movement pattern. Instead, we provide a one-stop solution (both in modeling and in technical)
for mining various co-movement patterns with flexible constraints. With
these constraints, users are able to define a more accurate pattern
which are free from the anomalies (e.g., loss-connection anomaly as in Figure 2).

Besides, the nature of GSP and GCMP are different. The goal of GSP 
is to find the sequences which conform to the gap constraint and occur frequently.
When counting the occurrence of a sequence in GSP, the object
which a sequence belongs to does not matter. Therefore, for
a pattern result in GSP, it cannot guarantee either the 
distinction of involved objects or the closeness of those objects.
In contrast, GCMP has requirements on both the spatial closeness 
of objects and the number of distinct objects.
Due to the unawareness of objects, techniques that are used in GSP 
cannot be directly applied on GCMP. We also summarize the difference
in Section 2.5.
%
%When counting the occurrence of a sequence, 
%the object the sequence belongs does not matter. 
%Therefore, a pattern in FS may corresponds to any objects where
%the closeness or distinction of objects are not statisfied. Due
%to this difference, FS solutions cannot be directly applied on GCMP.
%
%
%We agree with the reviewer that gap parameter is 
%also used in the Sequence Mining problem. However, 
%Sequence mining problem is different from the GCMP mining problem. 
%The major difference is that in sequence mining problem, the object
%itself is not considered. In particular, the frequency of a sequence is 
%counted as the occurrence of a sequence. Thus, an object may
%contribute multiple occurrence of a sequence. Conversely, GCMP
%in addition to impose the gap parameter in the temporal constraint, also 
%require the objects in the pattern to be close. Due to this difference,
%Sequence mining solutions cannot support GCMP mining.
%
%that the gap parameter is an old concept. 
%However, the novelty of GCMP lies in the unification and generalization of existing patterns which provide
%a one-stop solution for co-movement pattern detection. And the
%gap parameter merely is used for solving the loose-connection anomaly (as in Figure 2). 
%
%Moreover, the problem of Sequence mining is also different from Trajectory mining.
%The major difference is that sequence mining do not consider the distinction between
%objects and the closeness among objects. This makes the solutions in Sequence mining
%unable to be applied in GCMP. We summarize these points in Section 2.5 of the revised paper.
%
%We agree the fact that the gap parameter is an old concept. Nevertheless, our major contribution in
%modeling lies in unifying existing movement patterns in the trajectory domain. 
%The gap parameter is naturally introduced to 
%solve the loose-connection anomaly (as in Figure 2).
%%which is suffered by the state-of-the-art generalization (i.e., Platoon pattern). 
%In the revision, we add Section 2.5 to accredit the origin of the \emph{gap} concept
%and distinguish the sequential mining problem with our problem. 
}

@@

\response{We appreciate the reviewer's concerns on the 
size of the output. 
In fact, both TRPM and SPARE
have considered compressing the output by only discovering 
the patterns with larger object set. 
Particularly,  in the Line Sweeping Mining (Algorithm 1) of TRPM, whenever
a pattern $c$ becomes valid, it is directly outputted (Lines 14-16). This
prohibits outputting $c$'s subset. Similarly, in the Apriori Enumerator (Algorithm 3)
of SPARE, a pattern $c$ is outputted if non of its superset would become a valid pattern (Line 12-14).
This prohibits the output $c$ to be a subset of any patterns.
These mechanisms effectively condense the output size by subsuming smaller
patterns using their supersets.
% Nevertheless, these mechanisms are able to preserve ALL valid patterns.
Linking to the Taxi example, if the ground truth contains 1000 taxis travel
together, then both TRPM and SPARE tend to output the single pattern.
In addition, real detection of ``anomalous'' co-moment pattern needs to 
impose stringent constraints via $M,K,L,G$. This would further reduces 
the number of potential patterns.
%On the other hand, in reality, if users wish to detect ``anomalous'' patterns,
%the parameters $M,K,L,G$ will be set to be very stringent. This reduces
%the size of potential outputs. For example, when detecting the Taxi's movements
%without ground truth of 1000 taxis, an input could be $M = 100$, $K=50$ min, $L=30$ min,
%$G = 2$ min. Then, there would be very less patterns satisfying these constraints.
% and the meaningfulness of the output. 
%In terms of the output size, 
%we certainly do not need to output every pattern.
%In fact, both TRPM and SPARE algorithms tend to only output
%the patterns with larger object size. That is to say, for
%the Taxi example, the pattern containing all the 1000 taxis will
%be outputted. To see this, in Algorithm 1, TRPM effectively 
%starts with a temporally invalid pattern and continue
%to shrink its object set until it is temporally valid. Then, the result
%is the ``largest'' valid pattern for these object. Similarly,
%in Algorithm 3, SPARE effectively starts with a temporally valid
%pattern with small object set. Then SPARE continues to grow
%the object set until its temporal dimension is about to be non-decomposable (Definition 5).
%Therefore both TRPM and SPARE tends to find the ``largest'' valid
%patterns.
%
%In terms of the meaningfulness, we believe that all patterns that conformed
%to the parameters need to be outputted. 
%This is because the parameters
%
%
%We see two questions derived from this concern. First, do we need
%to output all ${1000 \choose M}$ patterns if $1000$ taxis form a true GCMP
%pattern? Our answer to this question is no. Therefore, the nature of both
%of our TRPM and SPARE algorithms tend to only output the patterns containing 
%larger sets of objects. To see this, in Algorithm 1, TRPM starts with an temporal invalid
%pattern and continues to shrink its object set while growing its temporal dimension.
%TRPM stops as soon as the object become valid. In an inverse way, in Algorithm 3, 
%SPARE starts with a pattern with small objects and continues to grow its object set
%while reducing its temporal dimension. SPARE stops as soon as the pattern is about
%to be temporally invalid. Therefore, both TRPM and SPARE would prefer to output
%larger pattern.
%
%
%%
%%We agree with the reviewer that if no other parameters are given, the maximum number
%%of patterns discovered could be as large as ${ \mathbb{O} \choose M}$, which may
%%be less interesting to users.
%However, such a case rarely occurs due to two reasons. First,
%in real trajectory mining, parameter $M$
%is associated with temporal parameters $K,L,G$ and they jointly prune many false
%patterns. Second, the nature of our solutions prefer 
%a larger pattern when possible~\footnote{TRPM shrinks an invalid larger pattern until it is valid. SPARE grows a valid smaller pattern until it is about to be invalid. Therefore both algorithms tend to find larger patterns.}.
%For example, if $(o_1,o_2,o_3)$ is 
%a pattern and $M = 2$, 
%then both TRPM and SPARE only output $(o_1,o_2,o_3)$ but not
%any of its subsets. This effectively compresses smaller patterns.
%
%In addition, when the temporal parameters $K,L,G$ are given,
%a larger $M$ actually decreases the number of resulted patterns. 
%This is because there are less patterns having more than $M$ objects. 
%As shown in Figures 7 (a)(g)(m), a larger $M$ actually reduces the running time of both TRPM and SPARE.
}

@@

\response{We agree that domain-specific constraints
on spatiotemporal mining could be similar to GCMP.
However, the domain-specific constraints (such as directions in road networks)
are nontrivial to be generalized. Since our GCMP model and technical solutions
do not leverage the domain knowledge, we can support
pattern discovery in broader scenarios such as user check-in history in social networks, visitor movements in a building and 
passenger flows in a city. We also add the discussion in Section 2.3.
}

### 2016-08-29 ReviewComments/review3.tex
Revised response
some more responses

\response{
We agree that leveraging more advanced Spark features
could further improve the performance of our solutions. 
We do not heavily leverage these features because 
we wish to focus on designing parallel GCMP mining algorithms.
%to provide a general parallel framework which does not tie to Spark.
Nevertheless, we are in progress in developing
a better open-sourced version with Spark-tailored optimization.

In the revision, we try to add a DataFrame based TRPM solution 
which aims to leverage the window functions from Spark-SQL.
However, to the best of our efforts, we could not achieve this goal in a short time. 
The major challenge is that current Spark do not support User Defined Aggregate Function (UDAF)
in window functions~\footnote{JIRA Spakr-8641: Native Spark Window Functions \url{https://issues.apache.org/jira/browse/SPARK-8641}}. Without UDAF, implementing
our Line sweep algorithm (Algorithm 1) using Spark primitives (e.g., sum, avg, rank and nth-tile)
is nontrivial. Nonetheless, we expect less performance
boost from TRPM even with the UDAF support. This is because the Line sweep algorithm is not
properly reducible, where Spark system would fail to leverage the partial aggregates
across multiple windows. As a result, it would be equivalent to our current TRPM
solution where windows are processed independently in parallel.
%TRPM although can be modeled using window functions over data frame, the aggregates
%it uses (i.e., the Line Sweeping method in Algorithm 1) cannot be simply represented
%by Spark primitives (e.g., sum, avg, rank, nth-tile). We then resort to
%the Spark User Defined Aggregate Function (UDAF). However, in Spark 1.5.2, the UDAF
%does not support window function syntax. Even in the latest Spark 1.6.0 (released 
%last month), UDAF does not fit with window function. We confirm these information
%with Spark development team~\footnote{In Spark JIRA tickets, fully supporting UDAF is still under development}.
%
%Nevertheless, since our Algorithm 1 is not reducible, even with UDAF, TRPM could
%not boost too much 
%
%we expect
%even with UDAF support TRPM would not boost too much.  Spark system is 
%%hard to leverage the partial aggregates acrocess
%
%we can analytically expect even using UDAF, TRPM would
%not boost too much. This is because our Algorithm 1 is not reducible. That
%is we cannot merge or share partial aggregates across different windows. In summary,
%We wish to convince the reviewer that current TRPM implementation is reasonable. 
}

>>>>

\response{
We thank the reviewer for suggesting a better way of implementing 
the baseline algorithm (i.e., TRPM). However, this could
hardly be done after our investigations. The major reason is
that current version of Spark SQL (i.e., 1.6.2 released in June)
does not support the User Defined Aggregate Function (UDAF) on window
function~\footnote{JIRA Spakr-8641: Native Spark Window Functions \url{https://issues.apache.org/jira/browse/SPARK-8641}}. Without UDAF, implementing
our Line sweep algorithm (Algorithm 1) using Spark SQL 
primitive aggregates (e.g., sum, avg, rank and nth-tile)
is beyond our abilities.

Further, we believe the performance boost on TRPM from Spark SQL
is very limited. The reason is that our 
Line Sweeping Mining algorithm (Algorithm 1) is not a 
properly reducible function. That is, the result of a 
smaller window could not be used to compute the result of a larger
window. In such a case, Spark SQL has to process each
window independently in parallel, which is equivalent to our TRPM implementation.
%We agree that leveraging more advanced Spark features
%could further improve the performance of our solutions. 
%We do not heavily leverage these features because 
%we wish to focus on designing parallel GCMP mining algorithms.
%%to provide a general parallel framework which does not tie to Spark.
%Nevertheless, we are in progress in developing
%a better open-sourced version with Spark-tailored optimization. 
%In the revision, we try to add a DataFrame based TRPM solution 
%which aims to leverage the window functions from Spark-SQL.
%However, to the best of our efforts, we could not achieve this goal in a short time. 
%The major challenge is that current Spark do not support User Defined Aggregate Function (UDAF)
%in window functions~\footnote{JIRA Spakr-8641: Native Spark Window Functions \url{https://issues.apache.org/jira/browse/SPARK-8641}}. Without UDAF, implementing
%our Line sweep algorithm (Algorithm 1) using Spark primitives (e.g., sum, avg, rank and nth-tile)
%is nontrivial. Nonetheless, we expect less performance
%boost from TRPM even with the UDAF support. This is because the Line sweep algorithm is not
%properly reducible, where Spark system would fail to leverage the partial aggregates
%across multiple windows. As a result, it would be equivalent to our current TRPM
%solution where windows are processed independently in parallel.
%TRPM although can be modeled using window functions over data frame, the aggregates
%it uses (i.e., the Line Sweeping method in Algorithm 1) cannot be simply represented
%by Spark primitives (e.g., sum, avg, rank, nth-tile). We then resort to
%the Spark User Defined Aggregate Function (UDAF). However, in Spark 1.5.2, the UDAF
%does not support window function syntax. Even in the latest Spark 1.6.0 (released 
%last month), UDAF does not fit with window function. We confirm these information
%with Spark development team~\footnote{In Spark JIRA tickets, fully supporting UDAF is still under development}.
%
%Nevertheless, since our Algorithm 1 is not reducible, even with UDAF, TRPM could
%not boost too much 
%
%we expect
%even with UDAF support TRPM would not boost too much.  Spark system is 
%%hard to leverage the partial aggregates acrocess
%
%we can analytically expect even using UDAF, TRPM would
%not boost too much. This is because our Algorithm 1 is not reducible. That
%is we cannot merge or share partial aggregates across different windows. In summary,
%We wish to convince the reviewer that current TRPM implementation is reasonable. 
}
